---
title: "Data Structures in R"
author: "Tony Yao-Jen Kuo"
output:
  revealjs::revealjs_presentation:
    highlight: pygments
    reveal_options:
      slideNumber: true
      previewLinks: true
---

```{r include=FALSE}
knitr::opts_chunk$set(results = "hold")
```

# An overview

## Data structures...

>- collects scalars
>- can be indexing
>- can be slicing
>- are iterable

## We are gonna talk about 6 of them

>- vector
>- list
>- (optional)factor
>- data.frame
>- (optional)matrix
>- (optional)array

# Vectors

## Characteristics of a vector

>- element-wise operation
>- uniformed class
>- supports logical filtering

## Why is there always a `[1]` before printed scalar?

## Using c() to create vectors

```{r}
player_names <- c("Jeremy Lin", "Michael Jordan", "Shaquille O'Neal")
player_heights <- c(191, 198, 216)
player_weights <- c(91, 98, 148)
player_names
player_heights
player_weights
```

## Using `[INDEX]` indexing a value from vectors

```{r}
player_names[1]
player_names[2]
player_names[3]
player_names[length(player_names)] # in case we have a long vector
```

## Using `[c(INDICE)]` slicing values from vectors

```{r}
player_names[2:3]
player_names[c(1, 3)]
```

## What will happen if we set a NEGATIVE index?

```{r}
# Try it yourself
```

## Vectors are best known for its...

- Element-wise operation

```{r}
player_heights_m <- player_heights / 100
player_heights
player_heights_m
```

## Practices: Using vector operations for players' BMIs

```{r eval=FALSE}
player_bmis <- # ...
```

## Beware of the types

```{r}
# Name, height, weight, has_ring
mj <- c("Michael Jordan", 198, 98, TRUE)
mj
class(mj[1])
class(mj[2])
class(mj[3])
class(mj[4])
```

## How to generate vectors quickly

```{r}
11:21
seq(from = 11, to = 21)
seq(from = 11, to = 21, by = 2)
seq(from = 11, to = 21, length.out = 6)
rep(7, times = 7)
```

## Getting logical values

```{r}
player_heights <- c(191, 198, 216)
player_weights <- c(91, 98, 148)
player_bmis <- player_weights/(player_heights*0.01)**2
player_bmis > 30
```

## Logical filtering

```{r}
player_bmis[player_bmis > 30]
```

## Practices: finding odd numbers in `random_numbers`

```{r}
set.seed(87)
random_numbers <- sample(1:500, size = 100, replace = FALSE)
```

## Vector is iterable

```{r eval=FALSE}
for (ITERATOR in ITERABLE) {
  # do something iteratively until ITERATOR hits the end of ITERABLE
}
```

## Iterator as values

```{r}
player_heights <- c(191, 198, 216)
for (ph in player_heights) {
  print(ph*0.01)
}
```

## Not just printing it out...

```{r}
player_heights <- c(191, 198, 216)
player_heights_m <- c()
for (ph in player_heights) {
  player_heights_m <- c(player_heights_m, ph*0.01)
}
player_heights_m
```

## Practices: Applying `fizz_buzz()` on 1:100

- if input can be divided by 3, return "fizz"
- if input can be divided by 5, return "buzz"
- if input can be divided by 15, return "fizz buzz"
- otherwise, return input itself

```{r}
## [1] 1 2 "fizz" 4 "buzz" ... 14 "fizz buzz" 16 ... 99 "buzz"
```

## Iterators as indice

```{r}
player_names <- c("Jeremy Lin", "Michael Jordan", "Shaquille O'Neal")
player_heights <- c(191, 198, 216)
for (i in 1:length(player_names)) {
  player_height_m <- player_heights[i]/100
  print(sprintf("%s is %s meter tall", player_names[i], player_height_m))
}
```

## Practices: Is x a prime?

```{r eval=FALSE}
is_prime(87) ## FALSE
is_prime(89) ## TRUE
is_prime(91) ## FALSE
```

## Practices: How many primes are there between x and y?

```{r eval=FALSE}
count_primes(5, 11) ## 3
count_primes(5, 13) ## 4
count_primes(5, 15) ## 4
```

## Iterate with another style

```{r eval=FALSE}
while (CONDITION) {
  # do something iteratively when CONDITION == TRUE
}
```

## Iterators as indice

```{r}
i <- 1
while (i <= length(player_names)) {
  player_height_m <- player_heights[i]/100
  print(sprintf("%s is %s meter tall", player_names[i], player_height_m))
  i <- i + 1
}
```

## Practices: How many times do I have to flip a coin to get 6 heads?

## `for` is necessary condition for `while`

## Practices: Fibonacci

- Try using 2 types of loop to generate a certain fibonacci array.

```{r eval=FALSE}
fibonacci(0, 1, 5) ## [1] 0, 1, 1, 2, 3
fibonacci(0, 1, 7) ## [1] 0, 1, 1, 2, 3, 5, 8
fibonacci(0, 1, 9) ## [1] 0, 1, 1, 2, 3, 5, 8, 13, 21
```

## Practices: Poker card deck

```{r}
suits <- c("Spade", "Heart", "Diamond", "Clover")
ranks <- c("Ace", 2:10, "Jack", "Queen", "King")
```

# Lists

## Characteristics of lists

>- Different classes
>- Supports `$` selection like attributes

## Using `list()` to create a list

```{r}
infinity_war <- list(
  "Avengers: Infinity War",
  2018,
  8.6,
  c("Action", "Adventure", "Fantasy")
)
class(infinity_war)
```

## Check the apperance of a list

```{r}
infinity_war
```

## Using `[[INDEX]]` indexing list

```{r}
for (i in 1:length(infinity_war)) {
  print(infinity_war[[i]])
}
```

## Giving names to elements in list

```{r}
infinity_war <- list(
  movieTitle = "Avengers: Infinity War",
  releaseYear = 2018,
  rating = 8.6,
  genre = c("Action", "Adventure", "Fantasy")
)
infinity_war
```

## Using [["ELEMENT"]] indexing list

```{r}
for (e in names(infinity_war)) {
  print(infinity_war[[e]])
}
```

## Using `$ELEMENT` indexing list

```{r}
infinity_war$movieTitle
infinity_war$releaseYear
infinity_war$rating
infinity_war$genre
```

## Every element keeps its original class

```{r}
for (e in names(infinity_war)) {
  print(class(infinity_war[[e]]))
}
```

## Practices: Getting favorite players' last names in upper cases

Hint: using `strsplit()` to split players' name and using `toupper()` for upper cases.

```{r}
fav_players <- c("Steve Nash", "Paul Pierce", "Dirk Nowitzki", "Kevin Garnett", "Hakeem Olajuwon")
# [1] "NASH" "PIERCE" "NOWITZKI" "GARNETT" "OLAJUWON"
```

# (optional)Factors

## Characteristics of factors

>- Acts like a character vector
>- Unique character is recorded as **Levels**
>- Supports ordinal values and each character is encoded as **integers**
>- Default class of a character column

## Using `factor()` to create a factor

```{r}
all_time_fantasy <- c("Steve Nash", "Paul Pierce", "Dirk Nowitzki", "Kevin Garnett", "Hakeem Olajuwon")
class(all_time_fantasy)
all_time_fantasy <- factor(all_time_fantasy)
class(all_time_fantasy)
```

## Unique character in factor is recorded with levels

```{r}
rgbs <- factor(c("red", "green", "blue", "blue", "green", "green"))
rgbs
```

## Supports ordinal values

```{r}
temperatures <- factor(c("freezing", "cold", "cool", "warm", "hot"),
                       ordered = TRUE)
temperatures
temperatures[1] > temperatures[3]
```

## Adjusting the order of a factor

```{r}
temperatures <- factor(c("freezing", "cold", "cool", "warm", "hot"),
                       ordered = TRUE,
                       levels = c("freezing", "cold", "cool", "warm", "hot"))
temperatures
```

## Elements in factor are encoded as integers

```{r eval=FALSE}
temperatures <- c("freezing", "cold", "cool", "warm", "hot")
as.numeric(temperatures) # Error
temperatures <- factor(c("freezing", "cold", "cool", "warm", "hot"))
as.numeric(temperatures)
```

## Factors sometimes are hard to handle...

```{r}
all_time_fantasy <- factor(c("Steve Nash", "Paul Pierce", "Dirk Nowitzki", "Kevin Garnett", "Hakeem Olajuwon"))
all_time_fantasy <- c(all_time_fantasy, "Ray Allen")
all_time_fantasy
```

# Data Frames

## Characteristics of data frames

>- Has 2 dimensions `m x n` as in `rows x columns`
>- Rows are denoted as observations, while columns are denoted as variables
>- Each column has its own class
>- Supports `$` selection like attributes

## Using `data.frame()` to create a data frame

```{r}
player_names <- c("Jeremy Lin", "Michael Jordan", "Shaquille O'Neal")
player_heights <- c(191, 198, 216)
player_weights <- c(91, 98, 148)
has_rings <- c(FALSE, TRUE, TRUE)
player_df <- data.frame(player_names, player_heights, player_weights, has_rings)
```

```{r echo=FALSE}
knitr::kable(player_df)
```

## Character vectors are encoded as factors by default

```{r}
str(player_df)
```

## Using `stringsAsFactors = FALSE` for character class

```{r}
player_df <- data.frame(player_names, player_heights, player_weights, has_rings, stringsAsFactors = FALSE)
str(player_df)
```

## Selecting column from data frames as a vector

- Using column names in double quotes
- or column indice

```{r}
player_df[["player_names"]]
player_df[, "player_names"]
player_df[, 1]
```

## Or using `$` like attributes

```{r}
player_df$player_names
```

## Subsetting observations from data frames

- Using row indice

```{r}
player_df[c(2, 3), ]
```

## More commonly, using a logical vector

```{r}
player_df[player_df$has_rings, ]  # players with rings
player_df[!player_df$has_rings, ] # players without rings
```

## Creating logical vectors using operators

- Remember putting logical vector at the **row** index

```{r}
player_df$player_heights > 200
player_df[player_df$player_heights > 200, ]
```

# (Optional) Matrix

## Creating a matrix using `matrix()`

```{r}
my_mat <- matrix(1:4, nrow = 2)
class(my_mat)
```

## matrix operations

- Using `*` for element-wise multiplication
- Using `t` for transpose
- Using `%*%` for matrix multiplication

```{r}
my_mat <- matrix(1:4)
my_mat * my_mat
t(my_mat) %*% my_mat
```

## Practices: Make a 9 x 9 multiplication matrix

```{r echo=FALSE}
vec <- matrix(1:9)
vec %*% t(vec)
```

# (Optional) Array

## Using `array()` to create an array

```{r}
my_arr <- array(1:24, dim = c(4, 3, 2))
my_arr
```