---
title: "Getting Data with R"
author: "Tony Yao-Jen Kuo"
output:
  revealjs::revealjs_presentation:
    highlight: pygments
    reveal_options:
      slideNumber: true
      previewLinks: true
---

# How to get data with R

## Overview

>- From files
>- From web

# Getting data from files

## Using `read.csv()` for CSV files

- CSV stands for **comma separated values**

```{r}
file_url <- "https://storage.googleapis.com/ds_data_import/chicago_bulls_1995_1996.csv"
df <- read.csv(file_url, stringsAsFactors = FALSE) # personal preference...
dim(df)
```

## Using `read.table()` for general tabular text files

```{r}
file_url <- "https://storage.googleapis.com/ds_data_import/chicago_bulls_1995_1996.txt"
df <- read.table(file_url, header = TRUE, sep = ";")
dim(df)
```

## Using `readxl` package for Excel spreadsheets

```{r eval=FALSE}
install.packages("readxl")
```

```{r}
library(readxl)

file_path <- "/Users/kuoyaojen/Downloads/fav_nba_teams.xlsx"
chicago_bulls <- read_excel(file_path)
head(chicago_bulls)
```

## Importing other sheets

```{r eval=FALSE}
boston_celtics <- read_excel(file_path, sheet = "boston_celtics_2007_2008")
head(boston_celtics)
```

## Reading specific cell ranges

```{r eval=FALSE}
partial_chi <- read_excel(file_path, range = "B8:C13", col_names = FALSE)
knitr::kable(partial_chi)
```

## Using `jsonlite` package for JSON files

```{r eval=FALSE}
install.packages("jsonlite")
```

```{r}
library(jsonlite)

file_url <- "https://storage.googleapis.com/ds_data_import/chicago_bulls_1995_1996.json"
chicago_bulls <- fromJSON(file_url)
class(chicago_bulls)
```

## A quick review

|Source|Format|
|------|------|
|CSV|`data.frame`|
|TXT|`data.frame`|
|Spreadsheet|`data.frame`|
|JSON|`list`|

# Getting data from web

## `jsonlite` for RESTful APIs

```{r}
library(jsonlite)

web_url <- "https://ecshweb.pchome.com.tw/search/v3.3/all/results?q=macbook&page=1&sort=rnk/dc"
macbook <- fromJSON(web_url)
class(macbook)
names(macbook)
```

## `rvest` for HTML documents

```{r eval=FALSE}
install.packages("rvest")
```

## The use of `%>%` operator

- Originated from `magrittr` package
- Now an important operator for the `tidyverse` eco-system
- Can be generated with: Ctrl + Shift + m

## How to call a function

```{r}
library(rvest)

# traditional
sum(1:10)

# using %>%
1:10 %>% 
  sum()
```

## More examples

```{r}
# traditional
toupper(paste0(strsplit("Jeremy Lin", split = " ")[[1]][2], "sanity"))

# using %>% 
"Jeremy Lin" %>% 
  strsplit(split = " ") %>% 
  `[[` (1) %>% 
  `[` (2) %>% 
  paste0("sanity") %>% 
  toupper()
```

## `read_html()` for reading all html contents

```{r}
library(rvest)

mi_url <- "https://www.imdb.com/title/tt4912910/"
html_doc <- mi_url %>% 
  read_html()
```

## `html_nodes()` to locate elements

```{r}
html_doc %>% 
  html_nodes("strong span") # CSS selector
```

## `html_text()` to remove tags

```{r}
html_doc %>% 
  html_nodes("strong span") %>% 
  html_text()
```

## Data of html document are characters

```{r}
html_doc %>% 
  html_nodes("strong span") %>% 
  html_text() %>% 
  as.numeric()
```

## How to locate elements?

>- By CSS Selectors
>- By XPath

# The use of Chrome plugins

## SelectorGadget

A Chrome plugin for CSS selectors:  [SelectorGadget](https://chrome.google.com/webstore/detail/selectorgadget/mhjhnkcfbdhnjickkkdbjoemdmbfginb)

## How to use SelectorGadget?

## XPath Helper

A Chrome plugin for XPath: [XPath Helper](https://chrome.google.com/webstore/detail/xpath-helper/hgimnogjllphhhkhlmebbmlgjoejdpjl)

## How to use XPath Helper?

## Practices: Getting genre information from IMDB.com

```{r}
mi_url <- "https://www.imdb.com/title/tt4912910/"
```

```{r echo=FALSE}
html_doc %>% 
  html_nodes(".subtext a") %>% 
  html_text() %>% 
  `[` (-4)
```

## Practices: Getting cast information from IMDB.com

```{r}
mi_url <- "https://www.imdb.com/title/tt4912910/"
```

```{r echo=FALSE}
html_doc %>% 
  html_nodes(".primary_photo+ td a") %>% 
  html_text()
```

## Practices: Getting price ranking from Yahoo! Stock

- Top 100 for TSE: <https://tw.stock.yahoo.com/d/i/rank.php?t=pri&e=tse&n=100>
- Top 100 for OTC: <https://tw.stock.yahoo.com/d/i/rank.php?t=pri&e=otc&n=100>