---
title: "Getting and Cleaning Data"
author: 郭耀仁
output: 
  revealjs::revealjs_presentation:
    theme: black
    highlight: zenburn
    center: true
---

# 案例來源

## Getting and cleaning data by Johns Hopkins University on Coursera

<https://www.coursera.org/learn/data-cleaning>

# 資料

## 下載資料

[UCI HAR Dataset](https://storage.googleapis.com/jhu_coursera_data/UCI_HAR_Dataset.zip)

## 資料描述

Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data is collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

## 步驟指引

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

## 參考解答

```{r echo=FALSE, message=FALSE}
# run_analysis.R
library(dplyr)

## get_files: Getting files into R
get_files <- function(directory) {
  file_paths <- c("train/X_train.txt", "train/y_train.txt", "train/subject_train.txt", "test/X_test.txt", "test/y_test.txt", "test/subject_test.txt", "features.txt")
  file_paths <- paste0(directory, file_paths)
  df_names <- c("X_train", "y_train", "subject_train", "X_test", "y_test", "subject_test", "features")
  df_list <- list()
  for (i in 1:length(file_paths)) {
    df_list[[i]] <- read.table(file_paths[i], stringsAsFactors = FALSE)
  }
  names(df_list) <- df_names
  return(df_list)
}

## merge_training_test: Merges the training and the test sets to create one data set
merge_training_test <- function(df_list) {
  X <- rbind(df_list$X_train, df_list$X_test)
  y <- rbind(df_list$y_train, df_list$y_test)
  subject <- rbind(df_list$subject_train, df_list$subject_test)
  df <- cbind(y, X)
  df <- cbind(subject, df)
  return(df)
}

## extract_mean_std: Extracts only the measurements on the mean and standard deviation for each measurement
extract_mean_std <- function(df, df_list) {
  feature_names <- df_list$features$V2
  has_mean_std <- feature_names %>%
    grepl(pattern = "mean()|std()")
  has_meanFreq <- feature_names %>%
    grepl(pattern = "meanFreq()")
  feature_selector <- has_mean_std & !has_meanFreq
  feature_selector <- c(TRUE, TRUE, feature_selector)
  mean_std_df <- df[feature_selector]
  return(mean_std_df)
}

## recode_activity: Uses descriptive activity names to name the activities in the data set
recode_activity <- function(mean_std_df) {
  activities <- c("WALKING", "WALKING_UPSTAIRS", "WALKING_DOWNSTAIRS", "SITTING", "STANDING", "LAYING")
  for (i in 1:length(activities)) {
    mean_std_df[, 2][mean_std_df[, 2] == i] <- activities[i]
  }
  return(mean_std_df)
}

## label_variable_names: Appropriately labels the data set with descriptive variable names.
label_variable_names <- function(mean_std_df, df_list) {
  feature_names <- df_list$features$V2
  has_mean_std <- feature_names %>%
    grepl(pattern = "mean()|std()")
  has_meanFreq <- feature_names %>%
    grepl(pattern = "meanFreq()")
  feature_selector <- has_mean_std & !has_meanFreq
  variable_names <- c("subject", "activity", feature_names[feature_selector])
  names(mean_std_df) <- variable_names
  return(mean_std_df)
}

## get_summarized_data: From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
get_summarized_data <- function(mean_std_df) {
  tidy_data <- mean_std_df %>%
    group_by(subject, activity) %>%
    summarise_all(mean)
  return(tidy_data)
}
```

```{r echo=FALSE}
unzip("~/Downloads/UCI_HAR_Dataset.zip", exdir = "~/Downloads/")
df_list <- get_files("~/Downloads/UCI HAR Dataset/")
df <- merge_training_test(df_list)
mean_std_df <- extract_mean_std(df, df_list)
mean_std_df <- recode_activity(mean_std_df)
mean_std_df <- label_variable_names(mean_std_df, df_list)
tidy_data <- get_summarized_data(mean_std_df)
```

```{r}
dim(tidy_data)
```

---

```{r}
head(tidy_data)
```

