---
title: "互動視覺化"
author: 郭耀仁
output: 
  revealjs::revealjs_presentation:
    theme: black
    highlight: zenburn
    center: true
---

# 概念

## 讓圖形與資料互動

- 改變圖形中的元素
- 連結多個圖形
- 與瀏覽者製造更深的連結

## 常見互動的圖形元素

- 對資料做篩選（filter）
- 對度量做更動（scaling）
- 資料提示（annotations）

## 對資料做篩選（filter）

```{r echo=FALSE, message=FALSE}
library(plotly)
library(gapminder)

plot_ly(data = gapminder, type = "scatter", x = ~gdpPercap, y = ~lifeExp, color = ~continent, text = ~country, mode = "markers", hoverinfo = "text", frame = ~year)
```

## 對度量做更動（scaling）

```{r message=FALSE, echo=FALSE}
plot_ly(data = subset(gapminder, subset = year == 2007), type = "scatter3d", x = ~gdpPercap, y = ~lifeExp, z = ~pop, color = ~continent)
```

## 資料提示（annotations）

```{r echo=FALSE}
plot_ly(data = subset(gapminder, subset = year == 2007), type = "scatter", x = ~gdpPercap, y = ~lifeExp, color = ~continent, text = ~country, mode = "markers", hoverinfo = "text")
```

# 工具

## R 語言使用者

- [`plotly` 套件](https://plot.ly/)
- [`shiny` 套件](https://shiny.rstudio.com/)

## 網頁工程師（撰寫 JavaScript）

- [`D3.js` 框架](https://d3js.org/)：終極的互動圖形解決方案

# Plotly

---

> Plotly for R is an interactive, browser-based charting library built on the open source javascript graphing library, plotly.js. It works entirely locally, through the HTML widgets framework.

## https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

## 新手的 Plotly

```{r message=FALSE}
# install.packages("plotly")
library(plotly) # Load package
library(gapminder) # Load data
library(dplyr)

gapminder_2007 <- gapminder %>%
  filter(year == 2007)
p <- plot_ly(data = gapminder_2007, x = ~gdpPercap, y = ~lifeExp) # Save plot as p
```

---

```{r message=FALSE, warning=FALSE}
p # Display plot
```

## 基礎的 `plotly` 函數

- `plot_ly()`
- `plot_geo()`
- `layout()`
- `add_trace()`
- `style()`

## 可以這樣做函數呼叫： 

```{r eval=FALSE}
p <- plot_ly(...) %>%
  add_trace(...) %>%
  add_layout(...) # Save plot as p
p # Display plot
```

## 從 Gallery 著手

https://plot.ly/r/

# Shiny

---

> Shiny is an R package that makes it easy to build interactive web apps straight from R. You can host standalone apps on a webpage or embed them in R Markdown documents or build dashboards. You can also extend your Shiny apps with CSS themes, htmlwidgets, and JavaScript actions.

## 註冊一個 shinyapps 帳號

- 免費帳號可以 host 5 個 Shiny App
- https://www.shinyapps.io/

## 連結帳號與 RStudio

- Account
- Tokens
- show
- Copy to clipboard

## 基本架構

- Your Shiny App Name
  - `ui.R`
  - `server.R`

## `ui.R`

```{r eval=FALSE}
library(shiny)

shinyUI(
  fluidPage(
    titlePanel("My First Shiny App"),
      sidebarLayout(
        sidebarPanel(
          checkboxGroupInput("continents",
                             "Select continents:",
                             c("Asia" = "Asia",
                               "Europe" = "Europe",
                               "Africa" = "Africa",
                               "Americas" = "Americas",
                               "Oceania" = "Oceania"),
                             selected = c("Asia", "Europe", "Africa", "Americas", "Oceania")
                             )
          ),
        mainPanel(
          plotOutput("scatterPlot")
          )
        )
    )
  )
```

## `server.R`

```{r eval=FALSE}
library(shiny)
library(dplyr)
library(gapminder)
library(ggplot2)

# Define server logic required to draw a histogram
shinyServer(function(input, output) {
  output$scatterPlot <- renderPlot({
    selected_continents <- input$continents
    gapminder_plot <- gapminder %>%
      filter(year == 2007) %>%
      filter(continent %in% selected_continents)
    ggplot(gapminder_plot, aes(x = gdpPercap, y = lifeExp, colour = continent, size = pop)) +
      geom_point()
  })
})
```

## 讓篩選變得更流暢

- 針對 `input$` 變數使用 `reactive()` 函數

## 加入 year 的 slider

## 發佈 Shiny App 並且分享你的連結