---
title: "Project: Programming with R"
author: "Tony Yao-Jen Kuo"
output:
  revealjs::revealjs_presentation:
    highlight: pygments
    reveal_options:
      slideNumber: true
      previewLinks: true
---

```{r include=FALSE}
knitr::opts_chunk$set(results = "hold", message = FALSE)
```

# Project Overview

## Project source

Assignment from [Programming with R](https://www.coursera.org/learn/r-programming)

## Write 3 functions to interact with data

>- `pollutantmean(directory, pollutant, id = 1:332)`

>- `complete(directory, id = 1:332)`

>- `corr(directory, threshold = 0)`

## Getting data

[specdata.zip](https://storage.googleapis.com/jhu_coursera_data/specdata.zip)

## How to download, unzip data with R?

- `download.file()` for downloading
- `unzip()` for unzipping

## About data

- 332 CSV files after unzipping
- Each CSV file has 4 variables

# Function 1

## Try to calculate the mean value of certain pollutant from different stations

`pollutantmean(directory, pollutant, id = 1:332)`

## Hints for function 1

- Set `na.rm = TRUE` in `mean()` if there are NAs

## Sample outputs

```{r echo=FALSE}
pollutantmean <- function(directory, pollutant, id = 1:332) {
  csv_filenames <- c()
  for (i in id) {
    if (nchar(i) == 1) {
      csv_filename <- paste0("00", i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    } else if (nchar(i) == 2) {
      csv_filename <- paste0("0", i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    } else {
      csv_filename <- paste0(i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    }
  }
  csv_lst <- list()
  for (i in 1:length(csv_filenames)) {
    csv_files_dir <- paste0(directory, "/", csv_filenames[i])
    csv_lst[[i]] <- read.csv(csv_files_dir, stringsAsFactors = FALSE)
  }
  df <- csv_lst[[1]]
  if (length(csv_lst) != 1) {
    for (i in 2:length(csv_lst)) {
      df <- rbind(df, csv_lst[[i]])
    }
  }
  filtered_vector <- df[, pollutant]
  ans <- mean(filtered_vector, na.rm = TRUE)
  return(ans)
}
```

```{r}
my_dir <- "/Users/kuoyaojen/Downloads/specdata"
pollutantmean(my_dir, "sulfate", 1:10)
pollutantmean(my_dir, "nitrate", 70:72)
pollutantmean(my_dir, "nitrate", 23)
```

# Function 2

## Try to calculate how many complete rows are in different CSV files

`complete(directory, id = 1:332)`

## Hints for function 2

- Use `complete.cases()` to get complete rows from a data frame

## Sample output 1

```{r echo = FALSE}
complete <- function(directory, id = 1:332) {
  csv_filenames <- c()
  for (i in id) {
    if (nchar(i) == 1) {
      csv_filename <- paste0("00", i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    } else if (nchar(i) == 2) {
      csv_filename <- paste0("0", i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    } else {
      csv_filename <- paste0(i, ".csv")
      csv_filenames <- c(csv_filenames, csv_filename)
    }
  }
  csv_lst <- list()
  for (i in 1:length(csv_filenames)) {
    csv_files_dir <- paste0(directory, "/", csv_filenames[i])
    csv_lst[[i]] <- read.csv(csv_files_dir, stringsAsFactors = FALSE)
  }
  df_id <- id
  nobs <- c()
  for (i in 1:length(csv_lst)) {
    n_complete_cases <- sum(complete.cases(csv_lst[[i]]))
    nobs <- c(nobs, n_complete_cases)
  }
  return_df <- data.frame(id = df_id, nobs = nobs)
  return(return_df)
}
```

```{r}
my_dir <- "/Users/kuoyaojen/Downloads/specdata"
complete(my_dir, 1)
complete(my_dir, c(2, 4, 8, 10, 12))
```

## Sample output 2

```{r}
complete(my_dir, 30:25)
complete(my_dir, 3)
```

# Function 3

## Try to calculate the correlation coefficient for CSV files, which have complete observations over `threshold`

`corr(directory, threshold = 0)`

## Hints for function 3

- Use `cor(x, y, use = "pairwise.complete.obs")` function for correlation coefficient

```{r echo = FALSE}
corr <- function(directory, threshold = 0) {
  csv_filenames <- list.files(directory)
  csv_directories <- paste0(directory, "/", csv_filenames)
  csv_filelist <- list()
  for (i in 1:length(csv_filenames)) {
    csv_filelist[[i]] <- read.csv(csv_directories[i])
  }
  nobs <- c()
  for (i in 1:length(csv_filelist)) {
    n_complete_cases <- sum(complete.cases(csv_filelist[[i]]))
    nobs <- c(nobs, n_complete_cases)
  }
  filter_vector <- nobs >= threshold
  if (sum(filter_vector) == 0) {
    cor_vector <- c()
    return(cor_vector)
  } else {
    filtered_list <- csv_filelist[filter_vector]
    cor_vector <- c()
    for (i in 1:length(filtered_list)) {
      cor_vector[i] <- cor(filtered_list[[i]]$sulfate, filtered_list[[i]]$nitrate, use = "pairwise.complete.obs")
    }
    cor_vector <- cor_vector[!is.na(cor_vector)]
    return(cor_vector)
  }
}
```

## Sample output 1

```{r}
my_dir <- "/Users/kuoyaojen/Downloads/specdata"
cr <- corr(my_dir, 150)
head(cr)
summary(cr)
```

## Sample output 2

```{r}
cr <- corr(my_dir, 400)
head(cr)
summary(cr)
```

## Sample output 3

```{r}
cr <- corr(my_dir, 5000)
summary(cr)
length(cr)
```

## Sample output 4

```{r}
cr <- corr(my_dir)
summary(cr)
length(cr)
```