---
title: "Variable Types in R"
author: "Tony Yao-Jen Kuo"
output:
  revealjs::revealjs_presentation:
    highlight: pygments
    reveal_options:
      slideNumber: true
      previewLinks: true
---

```{r include=FALSE}
knitr::opts_chunk$set(results = "hold")
```

# An overview

## We are gonna talk about 7 of them

>- numeric
>- (optional)integer
>- (optional)complex
>- character
>- logical
>- Date
>- POSIXct

## Using `class()` to examine variable types

```{r}
class(87)
class("Luke Skywalker")
class(TRUE)
```

# Numerics

## Using numeric as the primary digit type

```{r}
my_lucky_number <- 24
class(my_lucky_number) # numeric
class(2.4) # numeric
class(-8.7) # numeric
class(0) # numeric
```

## Using L to denote the integer type

```{r}
my_lucky_integer <- 87L
class(my_lucky_integer) # integer
my_lucky_integer
```

## Using i to denote the imaginary part

```{r}
my_lucky_complex <- 8 + 7i
class(my_lucky_complex) # complex
my_lucky_complex
```

## Numeric operations

|Operators|Usage|
|---------|-----|
|`+`|add|
|`-`|minus|
|`*`|multiply|
|`/`|divide|
|`**` or `^`|power|
|`%%`|modulo|
|`%/%`|quotient|

> - Using `()` to prioritize operations

## Practices: Jeremy Lin's BMI

$$BMI = \frac{weight_{kg}}{height_{m}^2}$$

```{r}
jeremyLin_height <- 191
jeremyLin_weight <- 91
```

## Practices: NBA players' BMIs

```{r}
steveNash_height <- 191
steveNash_weight <- 82
shaq_height <- 216
shaq_weight <- 148
jordan_height <- 198
jordan_weight <- 98
```

## We need FUNCTION to help us

```{r eval=FALSE}
MY_FUNCTION <- function(x, y, arg1, arg2, ...) {
  # using x, y, arg1, arg2 to get output
  return(OUTPUT)
}
```

## `get_bmi()` function

```{r}
get_bmi <- function(height, weight) {
  bmi <- weight/(height*0.01)**2
  return(bmi)
}
get_bmi(steveNash_height, steveNash_weight)
get_bmi(shaq_height, shaq_weight)
get_bmi(jordan_height, jordan_weight)
```

# character

## Using '' or "" for characters

```{r}
mj <- "Michael Jordan"
class(mj)
mj <- 'Michael Jordan'
class(mj)
```

## When to use '' or ""?

```{r eval=FALSE}
# Try to assign one of the greatest center in NBA history
Shaquille O'Neal
```

## Practices: What did Ross Geller say?

```{r eval=FALSE}
Let's put aside the fact that you "accidentally" pick up my grand mother's ring.
```

## Using `sprintf()` for string print with format

```{r}
jordan_BMI <- get_bmi(jordan_height, jordan_weight)
sprintf("Michael Jordan's BMI is %s", jordan_BMI)
```

## Common formats using `sprintf()`

- `%s`: pure text
- `%f`: float
- `%e`: scientific notation

```{r eval=FALSE}
?sprintf # this function is well-documented!
```

# logical

## Black or White, Head or Tail

```{r eval=FALSE}
class(TRUE)
class(FALSE)
class(True)  # error
class(False) # error
class(true)  # error
class(false) # error
```

## Logical operators

|Operators|Usage|
|---------|-----|
|`>`|larger than|
|`>=`|larger than or equal to|
|`<`|smaller than|
|`<=`|smaller than or equal to|
|`==`|equal to|
|`!=`|not equal to|
|`%in%`|belong to|
|`!`|not|

## Comparing characters

`"Z" > "z" > "Y" > "y" > "X" > "x" ... > "B" > "b" > "A" > "a"`

## Connecting logicals

- `&` for logical AND
- `|` for logical OR

## Is Shaq overweight?

- Well, that depends on the definition of overweight

```{r}
# Overweight: BMI > 30 AND body fat > 25%

# Overweight: BMI > 30 OR body fat > 25%

```

## Using logicals in control statements

```{r eval=FALSE}
# 1 branch
if (CONDITION) {
  # do something when CONDITION equals TRUE
}
```

## Using `if-else` for 2 branches

```{r eval=FALSE}
# 2 branches
if (CONDITION) {
  # do something when CONDITION equals TRUE
} else {
  # do something when CONDITION equals FALSE
}
```

## Using `if-else if-else` for more than 3 branches

```{r eval=FALSE}
# 3 branches
if (CONDITION_1) {
  # do something when CONDITION_1 equals TRUE
} else if (CONDITION_2) {
  # do something when CONDITION_2 equals TRUE
} else {
  # do something when both CONDITION_1 and CONDITION_2 equals FALSE
}
```

## Practices: `get_bmi_label()`

<https://en.wikipedia.org/wiki/Body_mass_index>

```{r eval=FALSE}
get_bmi_label <- function(player_name, height, weight) {
  # ...
}
get_bmi_label("Jeremy Lin", 191, 91)
# [1] "Jeremy Lin's BMI label is Normal"
```

## Practices: `fizz_buzz()`

- if input can be divided by 3, return "fizz"
- if input can be divided by 5, return "buzz"
- if input can be divided by 15, return "fizz buzz"
- otherwise, return input itself

```{r eval=FALSE}
fizz_buzz <- function(x) {
  # ...
}
fizz_buzz(6)  # [1] "fizz"
fizz_buzz(10) # [1] "buzz"
fizz_buzz(30) # [1] "fizz buzz"
fizz_buzz(31) # [1] 31
```

# Class judgement and conversion

## Using `is.___()` for judgement

> - `is.numeric()`
> - `is.character()`
> - `is.logical()`

## Using `as.___()` for conversion

> - `as.numeric()`
> - `as.character()`
> - `as.logical()` 

# Date and Datetime

## Using `Sys.Date()` for current date

```{r}
Sys.Date()
```

## Mysterious number

```{r}
sys_date <- Sys.Date()
as.numeric(sys_date) # what is this number?
```

## Date originates from 1970-01-01

```{r}
sys_date - as.numeric(sys_date)
```

## Every integer stands for a specific date

```{r}
original_date <- sys_date - as.numeric(sys_date)
original_date - 1
original_date
original_date + 1
```

## Practices: The long last of Beatles

The Beatles were an English rock band formed in Liverpool in 1960. Using `Sys.Date()` and tell us for how many years, this band still gets tremendous popularity.

```{r}
beatles_formed <- 1960
# ...
```

## Using `Sys.time()` for current datetime

```{r}
Sys.time()
```

## Datetime originates from 1970-01-01 08:00:00

```{r}
sys_datetime <- Sys.time()
original_datetime <- sys_datetime - as.numeric(sys_datetime)
original_datetime
```

## Every integer stands for a specific second

```{r}
sys_datetime <- Sys.time()
original_datetime <- sys_datetime - as.numeric(sys_datetime)
original_datetime - 1
original_datetime
original_datetime + 1
```

## Using `OlsonNames()` for specific timezones

## Practices: The 921 earthquake

1999-09-21 01:47:16 the Jiji earthquake occured in Nantou, Taiwan with a Richter scale of 7.3. The first major after-shock occured on 1999-09-21 01:57:15. Let us know how long was it between these two shocks.